humanCdGF_predict_cfu

### 2016-10-11 code/HumanCdGF_predict_cfu.Rmd
add prediction of steadystate CFU from day 0

```{r predict median CFU, fig.width=6,fig.height=6}

@@

>>>>

```{r predict median CFU heatmap, fig.width=6,fig.height=6}

@@

### Reframing the question for predicting CFU
#### 1. Can Day 0 community predict the median steady-state-colonized CFU?
#### 2. Can Day 0 community predict the rate of colonization?

```{r predict median CFU, fig.width=10,fig.height=10}

Cage_dep_df <- meta_data %>% 
  filter(day == 0) %>% 
  select(human_source, sample_id, cage_id) %>% 
  left_join(rel_abund_shared_01)
Predict_cage_df <- select(Cage_dep_df, -human_source, -sample_id)
Predict_source_df <- select(Cage_dep_df, -cage_id, -sample_id)
Predict_cage_df$cage_id <- factor(Predict_cage_df$cage_id)
Predict_source_df$human_source <- factor(Predict_source_df$human_source)

# identify OTUs that predict cage/inocula
set.seed(seed)
rf_cage <- randomForest(cage_id ~ ., data=Predict_cage_df, importance=TRUE, ntree=n_trees)
set.seed(seed)
rf_source <- randomForest(human_source ~ ., data=Predict_source_df, importance=TRUE, ntree=n_trees)
features_source <- head(sort(rf_source$importance[ , 'MeanDecreaseAccuracy'], decreasing = T), 65)
features_cage <- head(sort(rf_cage$importance[ , 'MeanDecreaseAccuracy'], decreasing = T), 65)
cage_source_features <- unique(c(names(features_cage), names(features_source)))

colonized_cfu <- meta_data %>% 
  filter(day %in% c(4:10), !cage_id == 'NP1', 
         Early_Euth == FALSE, cdiff_strain == 431) %>% 
  select(mouse_id, cage_id, log_cfu) %>% 
  group_by(mouse_id, cage_id) %>% 
  summarise(log_cfu_med = median(log_cfu)) %>% 
  ungroup()
sample_list_d0 <- shared_list_days %>% 
    filter(day == 0) %>% 
    select(-day)
community_d0.df <- rel_abund_shared_01 %>% 
    right_join(sample_list_d0) %>% 
    select(-sample_id)
cfu_shared_colonized.df <- colonized_cfu %>% 
    right_join(community_d0.df) %>% 
    select(-mouse_id, -cage_id, -one_of(cage_source_features)) %>% 
    filter(!is.na(log_cfu_med))

set.seed(seed)
rf_all <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized.df,
                           importance = TRUE, ntree = n_trees, mtry = 14)

importance_sorted <- sort(importance(rf_all)[,1], decreasing = T)
top_important_OTU <- rownames(data.frame(importance_sorted[importance_sorted/max(importance_sorted) > 0.6]))

top_colonized_df <- data.frame(cfu_shared_colonized.df %>% 
  select(log_cfu_med, one_of(top_important_OTU)))

set.seed(seed)
rf_n_features <- randomForest(log_cfu_med ~ ., data=top_colonized_df,
                              importance=TRUE, ntree=n_trees, mtry = 2,
                              keep.forest=TRUE)

importance_final <- sort(importance(rf_n_features)[,1], decreasing = T)

rf_prediction <- data.frame(observed = cfu_shared_colonized.df$log_cfu_med, 
                       predicted = rf_n_features$predicted)
rf_prediction <- merge(rf_prediction, colonized_cfu, by.x = 'observed', by.y = 'log_cfu_med')

ggplot(data = rf_prediction, aes(x=observed, y=predicted, color = cage_id)) + geom_point() + 
  labs(x= 'Observed CFU (log10)', y = 'Predicted CFU (log10)',
       title = "Random Forest Prediction of Colonized C Difficile CFU from Day 0 Community") + 
  theme_bw() + xlim(0,9) + ylim(0,9) + 
  annotate('text', x=0.15*(max(rf_prediction$observed)), 
           y=0.95*(max(rf_prediction$predicted)), 
           label=paste0('R^2 = ', round(tail(rf_n_features$rsq, 1), 2)))

importance.df <- data.frame(otu = names(importance_final),
                            importance = importance_final,
                            tax = get_tax(1, names(importance_final), tax_file))
tax_label <- paste0(importance.df$tax, ' (', 
                    gsub('tu0*', 'TU ', importance.df$otu), 
                    ') : ', 
                    round(importance.df$importance, 3))

layout(matrix(c(1:9), ncol=3, byrow=TRUE))

cage_color <- rainbow(length(unique(rf_prediction$cage_id)))
for (i in seq_along(importance_final)) {
  partial_otu <- 
    partialPlot(rf_n_features, top_colonized_df, names(importance_final)[i],
                xlab='', main=paste(tax_label[i]), 
                ylim = c(min(top_colonized_df$log_cfu_med),max(top_colonized_df$log_cfu_med)))
  for(m in seq_along(unique(rf_prediction$cage_id))){
    cage <- unique(rf_prediction$cage_id)[m]
    points(y = top_colonized_df[rf_prediction$cage_id %in% cage, 'log_cfu_med'],
         x = top_colonized_df[rf_prediction$cage_id %in% cage , names(importance_final)[i]], 
         pch = 21, bg = cage_color[m])
  }
}
```
